### 2015-05-09 Math W1/math_w1_v2.tex
Imported initial Data
All already existing homework imported into Repo

\documentclass[fleqn,12pt, paper=a4,sffamily]{scrreprt}
\usepackage[utf8]{inputenc}
\usepackage[ngerman]{babel}
\usepackage{lastpage}
\usepackage{color}   %May be necessary if you want to color links
\usepackage{hyperref}
% for code samples
\usepackage{listings}
\usepackage{caption}
% font face
\usepackage{times}
% for drawings
\usepackage{tikz}
% math stuff
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% custom colors
\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}

\lstset{ %
language=Java,   							% choose the language of the code
basicstyle=\small\ttfamily,  				% the size of the fonts that are used for the code
numbers=left,                   			% where to put the line-numbers
numbersep=5pt,                  			% how far the line-numbers are from the code
backgroundcolor=\color{light-light-gray},   % choose the background color. You must add
frame=lrtb,           						% adds a frame around the code
tabsize=4,          						% sets default tabsize to 2 spaces
captionpos=b,           					% sets the caption-position to bottom
breaklines=true,        					% sets automatic line breaking
xleftmargin=1.5cm,							% space from the left paper edge
commentstyle=\color{pgreen},
keywordstyle=\color{pblue},
literate=%
    {Ö}{{\"O}}1
    {Ä}{{\"A}}1
    {Ü}{{\"U}}1
    {ß}{{\ss}}1
    {ü}{{\"u}}1
    {ä}{{\"a}}1
    {ö}{{\"o}}1
    {~}{{\textasciitilde}}1
}
\renewcommand{\lstlistingname}{Code}
\captionsetup[lstlisting]{font={footnotesize},margin=1.5cm,singlelinecheck=false } % removes "Listing 1: "
\definecolor{light-light-gray}{gray}{0.95}
\let\stdsection\section
\renewcommand\section{\stdsection}

\title{Hausaufgabenabgabe 16. April 2015}
\subtitle{Mathematik für Informatiker II: Analysis und lineare Algebra}
\author{\\
    \huge{Louis Kobras}\\\large{6658699}\\\large{4kobras@informatik.uni-hamburg.de} \\ \\
    \huge{Utz Pöhlmann}\\\large{6663579}\\\large{4poehlma@informatik.uni-hamburg.de}
    }
\date{}

\begin{document}
    \maketitle
    \newpage
    \section*{Aufgabenbereich 1}
        \subsection*{Aufgabe 1}
            \begin{equation*}
                \begin{split}
                    2x_1~+~x_2~+~x_3 &=1
                \end{split}
            \end{equation*}

\end{document}