### 2015-05-22 FGI/FGI W8/fgi_w8_v1.tex
started work and did task 7.3(?)

\documentclass[parskip,12pt,paper=a4,sffamily]{article}
%alternate documentclass:
%\documentclass[parskip,12pt,paper=a4,sffamily]{scrartl}
\usepackage[utf8]{inputenc}
\usepackage[ngerman]{babel}
\usepackage{lastpage}
\usepackage{color}   %May be necessary if you want to color links
\usepackage{hyperref}
% code snippets
\usepackage{listings}
% listing captions
\usepackage{caption}
% font: times new roman
\usepackage{times}
% tikz being tikz
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{pgf}
% import math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
% contradiction lightning
\usepackage{stmaryrd}
% alignment options
\usepackage{ragged2e}
% page margins
\usepackage[margin=2.5cm]{geometry}

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}

\lstset{ %
language=Java,   							% choose the language of the code
basicstyle=\small\ttfamily,  				% the size of the fonts that are used for the code
numbers=left,                   			% where to put the line-numbers
numbersep=5pt,                  			% how far the line-numbers are from the code
backgroundcolor=\color{light-light-gray},   % choose the background color. You must add
frame=lrtb,           						% adds a frame around the code
tabsize=4,          						% sets default tabsize to 2 spaces
captionpos=b,           					% sets the caption-position to bottom
breaklines=true,        					% sets automatic line breaking
xleftmargin=1.5cm,							% space from the left paper edge
commentstyle=\color{pgreen},
keywordstyle=\color{pblue},
literate=%
    {Ö}{{\"O}}1
    {Ä}{{\"A}}1
    {Ü}{{\"U}}1
    {ß}{{\ss}}1
    {ü}{{\"u}}1
    {ä}{{\"a}}1
    {ö}{{\"o}}1
    {~}{{\textasciitilde}}1
}
\renewcommand{\lstlistingname}{Code}
\captionsetup[lstlisting]{font={footnotesize},margin=1.5cm,singlelinecheck=false } % removes "Listing 1: "
\definecolor{light-light-gray}{gray}{0.95}
\let\stdsection\section
\renewcommand\section{\stdsection}

% add line break for subtitle (size: large)
\title{Formale Grundlagen der Informatik I
    \\\large{
        Abgabe der Hausaufgaben
    }
    \\\large{
    	Übungsgruppe 24 am \today
    }
}
\author{~\\
	\Large{Louis Kobras}\\
	\large{6658699}\\ %Matrikelnummer; wenn nicht für Uni, auskommentieren
	\large{4kobras@informatik.uni-hamburg.de}\\
	\\
	\Large{Utz Pöhlmann}\\
	\large{6663579}\\
	\large{4poehlma@informatik.uni-hamburg.de}\\
	\\
	\Large{}Philipp Quach\\
	\large{6706421}\\
	\large{4quach@informatik.uni-hamburg.de}\\
}

% leave empty for no date on title page
% comment for auto-generated date
\date{\today}

\begin{document}
	\maketitle
	\newpage
	\section*{Aufgabe 7.3}
	\label{sec:a7.3}
	\subsection*{Aufgabe 7.3.1}
	\label{ssec:a7.3.1}
	Sei $\mu$ in $L_1 \cap L_2$.
	Dann ist $\mu$ in $L_1$ oder $L_2$, aber auch in $P$.\\
	Sei $\mu$ in $L_1$.
	Dann ist $\mu$ in $P$, aber auch in $L_1 \cap L_2$.\\
	Sei $\mu$ in $L_2$.
	Dann ist $\mu$ in $P$, aber auch in $L_1 \cap L_2$.\\
	Daraus folgt, dass $L_1,L_2 \in P\Rightarrow L_1 \cap L_2 \in P.$\\
	\\
	Sei $L_1 \underline{\subset} P$. 
	Dann gilt: $L_1=P \setminus \overline{L_1}$.
	Daraus ergibt sich $\overline{L_1} = P \setminus L_1~~~\Rightarrow~\overline{L_1}\in P$.
	\subsection*{Aufgabe 7.3.2}
	\label{ssec:a7.3.2}
	Sei $\mu$ in $L_1 \cap L_2$.
	Dann ist $\mu$ in $L_1$ oder $L_2$, aber auch $NP$.\\
	Sei $\mu$ in $L_1$.
	Dann ist $\mu$ in $NP$, aber auch in $L_1 \cap L_2$.\\
	Sei $\mu$ in $L_2$.
	Dann ist $\mu$ in $NP$, aber auch in $L_1 \cap L_2$.\\
	Daraus folgt, dass $L_1,L_2 \in NP\Rightarrow L_1 \cap L_2 \in NP.$\\
	%-------------------------------------------------%
	\section*{Aufgabe 7.4}
	\label{ssec:a7.4}
	%-------------------------------------------------%
	\section*{Aufgabe 7.5}
	\label{sec:a7.5}
	\subsection*{Aufgabe 7.5.0}
	\label{ssec:a7.5.0}
	\subsection*{Aufgabe 7.5.1}
	\label{ssec:a7.5.1}
	\subsection*{Aufgabe 7.5.2}
	\label{ssec:a7.5.2}
	%-------------------------------------------------%
	\section*{Aufgabe 7.6}
	\label{sec:a7.6}
\end{document}