### 2015-06-02 Krams für SX/latex101.tex
Started work on LaTeX script

\documentclass[parskip,12pt,paper=a4,sffamily]{article}
%alternate documentclass:
%\documentclass[parskip,12pt,paper=a4,sffamily]{scrartl}
\usepackage[utf8]{inputenc}
\usepackage[ngerman]{babel}
\usepackage{lastpage}
\usepackage{color}   %May be necessary if you want to color links
\usepackage{hyperref}
% code snippets
\usepackage{listings}
% listing captions
\usepackage{caption}
% font: times new roman
%\usepackage{times}
% tikz being tikz
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{pgf}
% import math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
% contradiction lightning
\usepackage{stmaryrd}
% alignment options
\usepackage{ragged2e}
% page margins
\usepackage[margin=2.5cm]{geometry}

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}

\lstset{ %
language=Tex,   							% choose the language of the code
basicstyle=\small\ttfamily,  				% the size of the fonts that are used for the code
numbers=left,                   			% where to put the line-numbers
numbersep=5pt,                  			% how far the line-numbers are from the code
backgroundcolor=\color{light-light-gray},   % choose the background color. You must add
frame=lrtb,           						% adds a frame around the code
tabsize=4,          						% sets default tabsize to 2 spaces
captionpos=b,           					% sets the caption-position to bottom
breaklines=true,        					% sets automatic line breaking
xleftmargin=1.5cm,							% space from the left paper edge
commentstyle=\color{pgreen},
keywordstyle=\color{pblue},
literate=%
    {Ö}{{\"O}}1
    {Ä}{{\"A}}1
    {Ü}{{\"U}}1
    {ß}{{\ss}}1
    {ü}{{\"u}}1
    {ä}{{\"a}}1
    {ö}{{\"o}}1
    {~}{{\textasciitilde}}1
}
\renewcommand{\lstlistingname}{Code}
\captionsetup[lstlisting]{font={footnotesize},margin=1.5cm,singlelinecheck=false } % removes "Listing 1: "
\definecolor{light-light-gray}{gray}{0.95}
\let\stdsection\section
\renewcommand\section{\stdsection}

% add line break for subtitle (size: large)
\title{LaTeX 101
    \\\large{
        Warum Informatiker darunter etwas Anderes verstehen als Menschen
    }
}
\author{~\\
	\Large{Louis Kobras}\\
	\large{4kobras@informatik.uni-hamburg.de}\\
	\\
	\Large{Hauke Stieler}\\
	\large{4stieler@informatik.uni-hamburg.de}
}

% leave empty for no date on title page
% comment for auto-generated date
\date{\today}

\begin{document}
	\maketitle
	\newpage
	\tableofcontents
	\clearpage
	\section{Einführung}
	\label{sec:intro}
	\section{Dokumentklassen}
	\label{sec:class}
	\subsection{Beamer}
	\label{ssec:beamer}
	\subsection{Article}
	\label{ssec:article}
	\section{Pakete}
	\label{ssec:package}
	\section{Ein Quellenverzeichnis erstellen}
	\label{sec:ref}
	\section{Mathmode}
	\label{sec:math}
	\section{TikZ}
	\label{sec:tikz}
	Optional, verhandelbar
	\section{Tools}
	\label{sec:tools}
	\subsection{Detexify}
	\label{ssec:detex}
	\subsection{TeXmaker}
	\label{ssec:texmaker}
\end{document}