### 2015-10-24 FS3/GDB/G11B01_Kobras-Pöhlmann-DerDritte.tex
nun aber der GDB Grundzettel

\documentclass[ngerman]{gdb-aufgabenblatt}

\renewcommand{\Aufgabenblatt}{1}
\renewcommand{\Ausgabedatum}{Mi. 14.10.2015}
\renewcommand{\Abgabedatum}{Fr. 30.10.2015}
\renewcommand{\Gruppe}{Kobras, Pöhlmann, DerDritte}
\renewcommand{\STiNEGruppe}{11}
\renewcommand{\Semester}{WS 2015/16}

\begin{document}

%------------------------------------------------
\section{Informationssysteme}
%------------------------------------------------

*Lösung*

%------------------------------------------------

*Mehr Lösung*

%------------------------------------------------
\section{Miniwelt}
%------------------------------------------------

*Lösung*

%------------------------------------------------

*Mehr Lösung*

%------------------------------------------------
\section{Transaktionen}
%------------------------------------------------

*Lösung*

%------------------------------------------------

*Mehr Lösung*

%------------------------------------------------
\section{Warm-Up MySQL}
%------------------------------------------------

*Lösung*

%------------------------------------------------

*Mehr Lösung*

\end{document}