### 2016-08-16 FS4/SEM/release/images/4seasons.rnw
SEM:präsi stuff

11 
    de.renew.gui.CPNDrawing 17 
        de.renew.gui.TransitionFigure "attributes" "attributes" 2 "FigureWithID" "Int" 1 "FillColor" "Color" 255 0 0 255 107 60 24 16 NULL  
        de.renew.gui.TransitionFigure "attributes" "attributes" 2 "FigureWithID" "Int" 2 "FillColor" "Color" 255 0 0 255 107 60 24 16 NULL  
        de.renew.gui.TransitionFigure "attributes" "attributes" 2 "FigureWithID" "Int" 3 "FillColor" "Color" 255 0 0 255 267 61 24 16 NULL  
        de.renew.gui.TransitionFigure "attributes" "attributes" 2 "FigureWithID" "Int" 4 "FillColor" "Color" 255 0 0 255 268 196 24 16 NULL  
        de.renew.gui.TransitionFigure "attributes" "attributes" 2 "FigureWithID" "Int" 5 "FillColor" "Color" 255 0 0 255 108 193 24 16 NULL  
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 6 "FillColor" "Color" 255 0 0 255 2 131 68 186 68 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.standard.ChopBoxConnector REF 2  
            CH.ifa.draw.figures.ChopEllipseConnector 
                de.renew.gui.PlaceFigure "attributes" "attributes" 2 "FigureWithID" "Int" 7 "FillColor" "Color" 255 0 0 255 185 58 20 20 NULL    REF 10 
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 8 "FillColor" "Color" 255 0 0 255 2 205 68 267 68 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.figures.ChopEllipseConnector REF 10  
            CH.ifa.draw.standard.ChopBoxConnector REF 3   
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 9 "FillColor" "Color" 255 0 0 255 2 278 77 278 134 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.standard.ChopBoxConnector REF 3  
            CH.ifa.draw.figures.ChopEllipseConnector 
                de.renew.gui.PlaceFigure "attributes" "attributes" 2 "FigureWithID" "Int" 10 "FillColor" "Color" 255 0 0 255 268 133 20 20 NULL    REF 19 
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 11 "FillColor" "Color" 255 0 0 255 2 278 153 279 196 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.figures.ChopEllipseConnector REF 19  
            CH.ifa.draw.standard.ChopBoxConnector REF 4   
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 12 "FillColor" "Color" 255 0 0 255 2 268 203 213 203 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.standard.ChopBoxConnector REF 4  
            CH.ifa.draw.figures.ChopEllipseConnector 
                de.renew.gui.PlaceFigure "attributes" "attributes" 2 "FigureWithID" "Int" 13 "FillColor" "Color" 255 0 0 255 193 193 20 20 NULL    REF 28 
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 14 "FillColor" "Color" 255 0 0 255 2 194 203 132 201 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.figures.ChopEllipseConnector REF 28  
            CH.ifa.draw.standard.ChopBoxConnector REF 5   
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 15 "FillColor" "Color" 255 0 0 255 2 119 193 119 145 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.standard.ChopBoxConnector REF 5  
            CH.ifa.draw.figures.ChopEllipseConnector 
                de.renew.gui.PlaceFigure "attributes" "attributes" 2 "FigureWithID" "Int" 16 "FillColor" "Color" 255 0 0 255 109 125 20 20 NULL    REF 37 
        de.renew.gui.ArcConnection "attributes" "attributes" 2 "FigureWithID" "Int" 17 "FillColor" "Color" 255 0 0 255 2 119 126 119 76 NULL 
            CH.ifa.draw.figures.ArrowTip 0.4 8.0 8.0 1  "CH.ifa.draw.figures.ArrowTip" 
            CH.ifa.draw.figures.ChopEllipseConnector REF 37  
            CH.ifa.draw.standard.ChopBoxConnector REF 2   NULL  29 176 539 600